---
title: "Solutions"
output: html_document
---

### Problem 1

a. 

```{r}
democracy <- read.csv(file = "democracy.csv", stringsAsFactors = FALSE)

table(democracy$CIVLIB)
table(democracy$POLLIB)

democracy <- read.csv(file = "democracy.csv", stringsAsFactors = FALSE, na.strings = ".")
```

b.

```{r, results='hide'}

library(dplyr)
library(tidyr)
library(knitr)

dem_by_variable <- 
  democracy %>%
  gather(variable, value, -COUNTRY, -YEAR, -REGION, -CTYNAME)
dem_by_variable

dem_summary_stats <- dem_by_variable %>%
  group_by(variable, CTYNAME) %>%
  summarise(mean = mean(value),
            median = median(value)) %>%
  gather(stat, value, -CTYNAME, -variable) %>%
  ungroup() %>%
  unite(variable_stat, variable, stat) %>%
  spread(variable_stat, value)
```

c.

```{r, results = 'asis'}
knitr::kable(dem_summary_stats)
```

d.

```{r}
library(ggplot2)
pollib_histogram <- ggplot(democracy, aes(x = POLLIB)) + geom_histogram()
```

e. 

```{r}
gdp_histogram <- ggplot(democracy, aes(x = GDPW)) + geom_histogram()
```

f. 

```{r}
gdp_log_histogram <- ggplot(democracy, aes(x = log(GDPW))) + geom_histogram()
```
g.

```{r}
lib_v_GDP <- ggplot(democracy, aes(x = POLLIB, y = GDPW)) + geom_boxplot(aes(group = POLLIB))
```

i.

```{r}
lib_v_logGDP <- ggplot(democracy, aes(x = POLLIB, y = log(GDPW))) + geom_point()
```

j. too many countries in non-oil countries?

```{r}

oil_boxplot <- ggplot(democracy, aes(x = factor(OIL), y = log(GDPW))) +
  geom_boxplot()

oil_boxplot
```

k.

```{r}
forty_cath <- democracy[democracy$CATH > 39, ]
mean(forty_cath$GDPW)
```

10295.13

l.

```{r}
sixty_plus <- democracy[democracy$ELF60 > .60, ]
mean(sixty_plus$GDPW, na.rm = TRUE) 
```

```{r}
sixty_below <- democracy[democracy$ELF60 < .60, ]
mean(sixty_below$GDPW, na.rm = TRUE)
```

```{r}
sixty_NA <- democracy[democracy$ELF60 < 0, ]
mean(sixty_NA$GDPW)
```
??

m. 

```{r}
ed_av <- democracy %>%
group_by(YEAR) %>%
summarize(ed_median = median(EDT, na.rm = TRUE))

ed_av_plot <- ggplot(ed_av, aes(x = YEAR, y = ed_median)) + geom_point() + geom_smooth(method = "lm")

ed_av_plot
```

o.

```{r}

yr_dem_av <- democracy %>%
group_by(YEAR, REG) %>%
summarize(ed_median = median(EDT, na.rm = TRUE))

yr_dem_plot <- ggplot(yr_dem_av, 
                      aes(x = YEAR, 
                          y = ed_median,
                          group = REG)) + 
  geom_point() + 
  geom_smooth(method = "lm", aes(colour = REG))

yr_dem_plot

```
n. ???

```{r}
democracy %>%
  filter(YEAR == 1985, !is.na(EDT)) %>%
  arrange(EDT) %>%
  slice(113/2)
```
q.

```{r}

percentiles <- 
  democracy %>%
  group_by(NEWC) %>%
  summarize(ELF60_p25 = quantile(ELF60, probs = .25, na.rm = T),
            ELF60_p75 = quantile(ELF60, probs = .75, na.rm = T))

kable(percentiles, format = "markdown")

```

### PROBLEM 2

```{r}
data("anscombe")
```
```{r message=FALSE, results='hide'}
library("dplyr")
library("tidyr")
anscombe2 <- anscombe %>%
	mutate(obs = row_number()) %>%
	gather(variable_dataset, value, - obs) %>%
	separate(variable_dataset, c("variable", "dataset"), sep = 1L) %>%
	spread(variable, value) %>%
	arrange(dataset, obs)

```

a.

```{r}
anscombe2 %>%
  group_by(dataset) %>%
  summarize(meanx = mean(x, na.rm = TRUE),
           meany = mean(y, na.rm = TRUE),
           sdx = sd(x, na.rm = TRUE),
           sdy = sd(y, na.rm = TRUE),
           cor_xy = cor(x, y))

anscombe2_reg1 <- lm(x ~ y, data = filter(anscombe2, dataset == 1))
anscombe2_reg1

anscombe2_reg2 <- lm(x ~ y, data = filter(anscombe2, dataset == 2))
anscombe2_reg2

anscombe2_reg3 <- lm(x ~ y, data = filter(anscombe2, dataset == 3))
anscombe2_reg3

anscombe2_reg4 <- lm(x ~ y, data = filter(anscombe2, dataset == 4))
anscombe2_reg4

```
You would expect them to be similar according to their coefficients.

b. why are all of these the same?

```{r}

ans_plot_1 <- ggplot(aes(x = x, y = y), 
                     data = filter(anscombe2, dataset == 1)) + 
  geom_point() +
  geom_smooth(method = "lm")
ans_plot_1


ans_plot_2 <- ggplot(aes(x = x, y = y), 
                     data = filter(anscombe2, dataset == 2)) + 
  geom_point() +
  geom_smooth(method = "lm")
ans_plot_2


ans_plot_3 <- ggplot(aes(x = x, y = y), 
                     data = filter(anscombe2, dataset == 3)) + 
  geom_point() +
  geom_smooth(method = "lm")
ans_plot_3


ans_plot_4 <- ggplot(aes(x = x, y = y), 
                     data = filter(anscombe2, dataset == 4)) + 
  geom_point() +
  geom_smooth(method = "lm")
ans_plot_4

```


### PROBLEM 3


```{r}
sprinters <- read.csv("sprinters.csv")
```

a.
```{r}
sprinters_orig <- sprinters[sprinters$year < 2005 & sprinters$olympics == 1, ]
```

b.
```{r}
reg1 <- lm(time ~ year + women, data = sprinters_orig)
```

For each increase in time, the year is increased by (-0.0126) which means that as time progresses, runners are getting faster. Not really sure what the "women" one means.

```{r}
reg2 <- lm(time ~ year * women, data = sprinters_orig)
```
This regression shows how year and women interact, which means that there is a negative relationship between time and women*year. So, over time, women are getting faster at (-0.0058)?

```{r}
reg3 <- lm(time ~ year, data = filter(sprinters_orig, women == 1))
```
This regression only includes women, and it finds that for each increase in year, the time decreases by (-0.0168).
```{r}
reg4 <- lm(time ~ year, data = filter(sprinters_orig, women == 0))
```
this regression only includes women, and it finds that for each increase in year, the time decreases by (-0.011), which is a slower rate than that for women.

c. 

```{r}

plot_data <- data.frame(actual = reg1$model$time, predicted = reg1$fitted.values)


regs_plot <- ggplot(plot_data, aes(x = actual, y = predicted)) + 
  geom_point() +
  geom_smooth(method = "lm") 

regs_plot
```

d. 
```{r}
new_db <- data.frame(women = c(1,0),
                     year = 2156)

predict(reg1, newdata = new_db)

```
a 7.775 sprint seems otherworldly and unrealistic. The progression of sprint times is not exactly linear, at some point it will plateau and this model will no longer work.


e.

```{r}

RSME_reg <- lm(time ~ year * women, data = sprinters)

mse <- mean(residuals(RSME_reg)^2)
rmse <- sqrt(mse)
rmse

olympics_db <- sprinters[sprinters$olympics == 1, ]

world_champ <- sprinters[sprinters$olympics == 0, ]

oly_reg <- lm(time ~ year * women, data = olympics_db)

wc_reg <- lm(time ~ year * women, data = world_champ)

mse_oly <- mean(residuals(oly_reg)^2)
rmse_oly <- sqrt(mse_oly)
rmse_oly

mse_wc <- mean(residuals(wc_reg)^2)
rmse_wc <- sqrt(mse_wc)
rmse_wc

```
predict_oly_db <- data.frame(olympics = c(1,0),
                     year = 2004)

predict(RSME_reg, newdata = predict_oly_db)
???

